---
title: "Breslow-Day Test"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
```

The notes for the topics on this page can be found in the lecture 19 folder on Canvas. 

# {.tabset .tabset-fade}

## R

To conduct a Brewlosw-Day test in R you can use `BreslowDayTest()` from the *DescTools* package. `BreslowDayTest()` requires a 2 x 2 x k table as its main argument; you can also specify an odds ratio for the null hypothesis (if not specified, the Mantel-Haenszal OR will be used). To create 2 x 2 x k tables from a data-set you can use `xtabs()`. 

### Example One

```{r}
library(DescTools)

teeth <- tibble(
  exposure = rep(c("tetra", "tetra", "notetra", "notetra"), 2), 
  discolor = rep(c("yes", "no"), 4), 
  age = c(rep("child", 4), rep("adult", 4)), 
  count = c(30, 25, 12, 17, 4, 6, 68, 94)
)

teeth_tabs <- xtabs(count ~ exposure + discolor + age, data = teeth)
teeth_tabs

BreslowDayTest(teeth_tabs)
```

### Example Two

```{r}
cervical <- tibble(
  smoke = rep(c("yes", "yes", "no", "no"), 2), 
  cancer = rep(c("yes", "no", "yes", "no"), 2), 
  partners = c(rep("zeroone", 4), rep("twomore", 4)), 
  count = c(12, 21, 25, 118, 96, 148, 92, 150)
)

cervical_tabs <- xtabs(count ~ smoke + cancer + partners, data = cervical)
cervical_tabs

BreslowDayTest(cervical_tabs)
```

## SAS

To conduct a Breslow-Day test in SAS you use the frequency procedure with the `cmh` argument specified in the options of the table statement. The 2 x 2 x k tables are built as usual, but with the following adjustment: `table k*exposure*disease`. 

### Example One

```{r, engine = "sas", engine.path = "C:/Program Files/SASHome/SASFoundation/9.4/sas.exe"}
title 'Teeth Discoloration Example';
  data teeth;    
  input exposure $ discolor $ age $ count;
  cards;
    tetra yes child 30 
    tetra no child 25
    notetra yes child 12 
    notetra no child 17
    tetra yes adult 4 
    tetra no adult 6 
    notetra yes adult 68 
    notetra no adult 94 
  ; 
run;
    
proc freq data = teeth order = data;
  table age*exposure*discolor / chisq relrisk cmh expected; 
  weight count; 
run; 
```

### Example Two

```{r, engine = "sas", engine.path = "C:/Program Files/SASHome/SASFoundation/9.4/sas.exe"}
title 'Cervical Cancer Example';
data cervical;
  input smoke $ cancer $ partners $ count;
  cards;
    yes yes zeorone 12
    yes no zeorone 21 
    no yes zeorone 25 
    no no zeorone 118 
    yes yes twomore 96 
    yes no twomore 148 
    no yes twomore 92 
    no no twomore 150 
    ; 
run;
    
proc freq data = cervical order = data;
  table partners*smoke*cancer / chisq relrisk cmh; 
  weight count; 
run; 
```



