---
title: "Chi-squared and Fisher's Exact"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

The notes for the topics on this page can be found in the lecture 8 folder on Canvas. 

# {.tabset .tabset-fade}

## R

To conduct a chi-squared test or Fisher's exact test in R you can use the `CrossTable()` function from the *descr* package. `CrossTable()` isn't set up to directly account for weighted cell counts. If you only have summary data you can first use `xtabs()` to build the contigency table and then use the `%>%` operator to 'pipe' the table into `CrossTable()`; the below examples illustrate this. For more help, type `?CrossTable` in the R console. 

```{r}
vietnam <- tibble(
  service = c(rep("yes", 2), rep("no", 2)), 
  sleep = c(rep(c("yes", "no"), 2)), 
  count = c(173, 160, 599, 851)
)

vietnam <- vietnam %>% 
  mutate(service = as_factor(service), 
         sleep = as_factor(sleep))
```

### Chi-squared 

To conduct a chi-squared test, set the `chisq` argument to `TRUE`.

```{r}
library(descr)

xtabs(count ~ service + sleep, data = vietnam) %>% 
  CrossTable(expected = TRUE, chisq = TRUE)
```

### Fisher's exact

To conduct a Fisher's exact test, set the `fisher` argument to `TRUE`

```{r}
xtabs(count ~ service + sleep, data = vietnam) %>% 
  CrossTable(fisher = TRUE)
```

## SAS

To conduct a chi-squared test or Fisher's exact test in SAS use the frequency procedure. The following code will produce return results for both a chi-squared test and a Fisher's exact test. The syntax is:

```{}
proc freq data = <dataset>; 
  table <row variable>*<column variable> / chisq expected; 
  weight <weighting variable if summary data>; 
run; 
```

### Chi-squared & Fisher's exact

```{r, engine = "sas", engine.path = "C:/Program Files/SASHome/SASFoundation/9.4/sas.exe"}
data vietnam; 
  input service $ sleep $ count; 
  cards; 
  yes yes 173
  yes no 160
  no yes 599
  no no 851
  ; 
run; 
  
title 'Chi-squared test example'; 
proc freq data = vietnam order = data; 
  table service*sleep / relrisk chisq expected; 
  weight count; 
run; 
```




