---
title: "Confidence Intervals for Single Proportions"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

The notes for the topics on this page can be found in the lecture 3 folder on Canvas. 

# {.tabset .tabset-fade}

## R

The `prop_test` function in the *catfun* package provides both approximated confidence intervals and exact confidence intervals for single proportions. For more information run `?prop_test` in the R console. Another possible function is `BinomCI()` in the *DescTools* package.

```{r warning=FALSE}
library(catfun)

prop_test(2, 6, method = "wald", conf.level = .95)
```

## SAS

To create confidence intervals for a single proportion in SAS use the frequency procedure. 

```{r, engine = "sas", engine.path = "C:/Program Files/SASHome/SASFoundation/9.4/sas.exe"}
data followup1; 
  input loss; 
  cards; 
  2
  4
  ;
run; 

title 'Loss to Follow-up Exact CI Example'; 
proc freq data = followup1 order = data; 
  table loss / binomial(p = 0.5) alpha = 0.05; 
  weight loss; 
run; 
```



