---
title: "Cochran-Mantel-Haenszel Test"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

The notes for the topics on this page can be found in the lecture 20 folder on Canvas. 

# {.tabset .tabset-fade}

## R

To conduct a Cochran-Mantel-Haenszel test in R you can use `mantelhaen.test()` from the *stats* package. In this example, we first create our dataset and then 2 x 2 x 2 table as shown elsewhere. The 2 x 2 x 2 table is then piped into `mantelhaen.test()`. Notice that you must set `correct = FALSE` as the default is to conduct a test with a continuity correction.

```{r}
library(tidyverse)
library(magrittr)

health_policy <- tibble(
  stress = rep(c("low", "low", "high", "high"), 2), 
  opinion = rep(c("favorable", "unfavorable"), 4), 
  region = c(rep("rural", 4), rep("urban", 4)), 
  count = c(55, 135, 7, 53, 48, 12, 96, 94)
) 

health_policy %<>%
  mutate(stress = fct_relevel(stress, c("low", "high")))

xtabs(count ~ stress + opinion + region, data = health_policy) %>%
  mantelhaen.test(correct = FALSE)
```

## SAS

To conduct a Cochran-Mantel-Haenszel test in SAS you use the frequency procedure with the `cmh` argument specified in the options of the `table` statement. 

```{r, engine = "sas", engine.path = "C:/Program Files/SASHome/SASFoundation/9.4/sas.exe"}
data healthpolicy;
  input stress $ opinion $ region $ count;
  cards;
    low favorable rural 55
    low unfavorable rural 135
    high favorable rural 7
    high unfavorable rural 53
    low favorable urban 48
    low unfavorable urban 12
    high favorable urban 96
    high unfavorable urban 94
    ;
run;
    
proc freq data = healthpolicy order = data;
  table region*stress*opinion / chisq relrisk cmh;
  weight count;
run;
```

