---
title: "Cochran-Armitage Trend Test"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

The notes for the topics on this page can be found in the lecture 23 folder on Canvas. 

# {.tabset .tabset-fade}

## R

To conduct a Cochran-Armitage Trend test in R you can use `CochranArmitageTest()` from the *DescTools* package. `CochranArmitageTest()` requires a frequency table as it's main argument. For more information and examples, run `?CochranArmitageTest`.

```{r}
library(tidyverse)
library(DescTools)

trend <- tibble(
  insomnia = c(rep("yes", 6), rep("no", 6)), 
  age_class = rep(seq(1:6), 2), 
  count = c(150, 250, 264, 302, 238, 176, 384, 496, 520, 403, 205, 123)
)

trend <- trend %>% 
  mutate(insomnia = fct_relevel(insomnia, c("yes", "no")))

trend_table <- xtabs(count ~ insomnia + age_class, data = trend)

CochranArmitageTest(trend_table)
```

## SAS

To conduct a Cochran-Armitage Trend test in SAS you use the frequency procedure with the `trend` argument specified in the options of the `table` statement. 

```{}
proc freq data = <dataset>; 
  table <row variable>*<column variable> / chisq trend; 
  weight <weighting variable if summary data>; 
run; 
```

```{r, engine = "sas", engine.path = "C:/Program Files/SASHome/SASFoundation/9.4/sas.exe"}
data trend; 
  input insomnia $ ageclass count @@; 
  cards; 
  yes 1 150 yes 2 250 yes 3 264 yes 4 302 yes 5 238 yes 6 176
  no 1 384 no 2 496 no 3 520 no 4 403 no 5 205 no 6 123
  ; 
run; 
  
proc freq data = trend order = data; 
  table insomnia*ageclass / chisq trend; 
  weight count; 
run;
```
