---
title: "Goodness of Fit"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
```

The notes for the topics on this page can be found in the lectures 16 and 17 folder on Canvas. 

# {.tabset .tabset-fade}

## R

To examine goodness of fit using deviance we will use `gof_deviance()` from *catfun*, to conduct a Hosmer-Lemeshow test we will use `hoslem.test()` from *ResourceSelection*. 

### Deviance

`gof_deviance()` takes two arguments: (1) a working logistic regression model and (2) a saturated working logistic regression model. 

```{r}
library(catfun)

cad <- tibble(
  gender = c(rep("male", 6), rep("female", 6)), 
  agecat = c(rep(c("<60", "60-69", "70+"), 4)), 
  event = c(rep(c(0, 0, 0, 1, 1, 1), 2)), 
  count = c(110, 131, 126, 13, 37, 45, 411, 110, 189, 41, 24, 12)
)

saturated <- glm(event ~ gender*agecat, weight = count, data = cad, family = "binomial")
working <- glm(event ~ gender + agecat, weight = count, data = cad, family = "binomial")

gof_deviance(working, saturated)
```

### Hosmer-Lemeshow

`hoslem.test()` requires three arguments: The first argument is a vector of the observed outcomes, the second argument is a vector of the predicted outcomes, and the third is the number of quintiles for conducting the test. 

```{r}
library(ResourceSelection)
library(readxl)

evans_county <- read_excel("C:/Users/niwi8/OneDrive/Documents/p8120_ta/data/evanscounty.xlsx")

evans_model <- glm(chd ~ cat + age + ecg + smk + chl + hpt, data = evans_county, family = "binomial")

hoslem.test(evans_model$y, fitted(evans_model), g = 10)
```

## SAS

### Deviance

To obtain goodness of fit based on deviance in SAS you can specify `scale = none` in the `model` argument. 

```{r, engine = "sas", engine.path = "C:/Program Files/SASHome/SASFoundation/9.4/sas.exe"}
data cad; 
  input gender $ agecat $ events trials; 
  cards; 
  male <60 13 123
  male 60-69 37 168
  male 70+ 45 171
  female <60 41 452
  female 60-69 24 134
  female 70+ 12 201
  ; 
run; 
            
proc logistic data = cad; 
  class gender (ref = 'female')
        agecat (ref = '<60') / param = ref; 
  model events / trials = gender agecat / cl scale = none; 
run; 
```

### Hosmer-Lemeshow

To conduct a Hosmer-Lemeshow test in SAS you specify `lackfit` in the `model` argument. 

```{r, engine = "sas", engine.path = "C:/Program Files/SASHome/SASFoundation/9.4/sas.exe"}
proc import datafile = 'C:/Users/niwi8/OneDrive/Documents/p8120_ta/data/evanscounty.xlsx'
            out = evanscounty
            dbms = xlsx; 
run; 
            
proc logistic data = evanscounty descending; 
  model chd = cat age ecg smk chl hpt / lackfit; run; 
```

