---
title: "Hypothesis Tests for a Single Proportion"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The notes for the topics on this page can be found in the lecture 4 folder on Canvas. 

# {.tabset .tabset-fade}

## R

One method to conduct a hypothesis test for a single proportion in R is to use the `prop_test` function in the *catfun* package. For more information run `?prop_test` in the R console. For further help, please watch the video [example](#r-demonstration).

### Score test

```{r warning = FALSE}
library(catfun)

prop_test(7, 50, method = "wald", p = 0.2)
```

### Exact test

The only thing that changes when conducting an exact test is that you must specify `exact = TRUE`

```{r warning = FALSE}
library(catfun)

prop_test(2, 6, method = "wald", p = 0.2, exact = TRUE)
```

### R demonstration

<div align="center">
<iframe width="854" height="480" src="https://www.youtube.com/embed/WblDls-luPk" frameborder="0" allowfullscreen></iframe>
</div>

## SAS

To conduct a hypothesis test for a single proportion in SAS you use the frequency procedure. For further help, please watch the video [example](#sas-demonstration).

The syntax is: 

```{}
proc freq data = <dataset>; 
  table <variable name> / binomial(p = <null value>) alpha = <alpha level>; 
  weight <weighting variable if summary data>; 
run; 
```

Variable name corresponds to the variable in the dataset, null value corresponds to your null hypothesis, alpha level corresponds to your pre-specified alpha level, and weight corresponds to the variable in the dataset that contains your counts (in these cases, the same variable that proceeds your table statement).

### Score test

```{r, engine = "sas", engine.path = "C:/Program Files/SASHome/SASFoundation/9.4/sas.exe"}
data followup; 
  input loss; 
  cards; 
  7
  43
  ;
run; 

title 'Loss to Follow-up Example'; 
proc freq data = followup order = data; 
  table loss / binomial(p = 0.2) alpha = 0.05; 
  weight loss; 
run; 
```

### Exact test

The only thing that changes when conducting an exact test is you must add `exact binomial` after the table statement.

```{r, engine = "sas", engine.path = "C:/Program Files/SASHome/SASFoundation/9.4/sas.exe"}
data followupEX;
  input loss; 
  cards; 
  2
  4
  ;
run;

title 'Loss to Follow-Up Exact Example';
proc freq data = followupEX order = data; 
  table loss / binomial(p = 0.2) alpha = 0.05;
  exact binomial;
  weight loss; 
run; 
```

### SAS demonstration

<div align="center">
<iframe width="854" height="480" src="https://www.youtube.com/embed/bgA5Xnj6zf4" frameborder="0" allowfullscreen></iframe>
</div>


