---
title: "Logistic Regression"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

The notes for the topics on this page can be found in the lectures 13, 14, and 15 folders on Canvas. 

# {.tabset .tabset-fade}

## R

To conduct logistic regression in R we can use the `glm` function with `family = binomial`. The general syntax for regression models in R is `response ~ covariate one + covariate two + ..., data = dataset`. It is useful to save your model as an object, you can then call other functions such as `summary` which will output your estimated coefficients. It is also useful to use `tidy` from the *broom* package to convert certain output from the object storing your model as a dataframe that can then be further manipulated.  

### Group level data

When working with aggregate or group level data you can simply specify the `weight` statement. In the following example, we "tidy" the output and then calculate the confidence interval for the exponentiated coefficient. You cannot have continuous predictors when working with summary data. 

```{r}
library(broom)

trial <- tibble(
  estrogen = c(rep(c("no", "yes"), 2)), 
  event = c(0, 0, 1, 1), 
  count = c(7479, 7755, 623, 751)
)

trial_mod <- glm(event ~ estrogen, weights = count, data = trial, 
                 family = "binomial")

summary(trial_mod)

broom::tidy(trial_mod) %>% 
  filter(term != "(Intercept)") %>% 
  transmute(odds_ratio = exp(estimate), 
         lower = exp(estimate - 1.96*std.error), 
         upper = exp(estimate + 1.96*std.error))
```

### Individual level data

If you are working with continuous predictors, it might be useful to examine a 5-unit increase in your variable instead of a single unit increase. To do this, we first "tidy" the model output and then use `transmute` to calculate this effect. 

```{r message = FALSE}
library(broom)

low_bwt <- read_csv("../../data/lowbwt.csv") %>% 
  mutate_at(vars("smoke", "visits", "low"), as_factor)

bwt_mod <- glm(low ~ age, data = low_bwt, family = "binomial")

summary(bwt_mod)

broom::tidy(bwt_mod) %>% 
  filter(term != "(Intercept)") %>% 
  transmute(odds_ratio = exp(estimate), 
         lower = exp(estimate - 1.96*std.error), 
         upper = exp(estimate + 1.96*std.error))

broom::tidy(bwt_mod) %>% 
  filter(term != "(Intercept)") %>% 
  transmute(or_5 = exp(estimate * 5), 
            lower = exp(estimate * 5 - 1.96*std.error*5), 
            upper = exp(estimate * 5 + 1.96*std.error*5))
```

### Likelihood ratio test

To conduct a likelihood ratio test for nested models in R you can use the `lrtest` function from the *lmtest* package. To do so, you first fit your two models and save them as objects. You can then call `lrtest(<insert model 1>, <insert model 2>)`. 

```{r message = FALSE}
library(lmtest)

bwt_mod_1 <- glm(low ~ smoke + age, data = low_bwt, family = "binomial")
bwt_mod_2 <- glm(low ~ smoke + age + visits, data = low_bwt, family = "binomial")

lrtest(bwt_mod_1, bwt_mod_2)
```

## SAS

To conduct logistic regression in SAS you can use `proc logistic`. Depending on which type of data layout you have, SAS code for fitting a logistic model will change. Some analyses of how well the model fits the data will depend on the type of data you have. We cannot compare the findings from a logistic model with a table analysis if we have subject level data that contains a continuous covariate (and we are interested in investigating that continuous covariate).

### Group level data

A single binary predictor: 

```{r, engine = "sas", engine.path = "C:/Program Files/SASHome/SASFoundation/9.4/sas.exe"}
title 'Example Estrogen Trial'; 
data trial; 
  input estrogen events trials; 
  cards; 
  0 623 8102
  1 751 8506
  ; 
run; 
  
proc format; 
  value estfmt 0 = 'placebo' 
               1 = 'estrogen'; 
run; 
               
proc logistic data = trial; 
  class estrogen (ref = 'placebo') / param = ref; 
  model events/trials = estrogen / cl; 
  format estrogen estfmt.; 
run; 
```

A multivariable example: 

```{r, engine = "sas", engine.path = "C:/Program Files/SASHome/SASFoundation/9.4/sas.exe"}
title 'Example Infection and Treatment adjusting for Hospital Risk';
data hosp;
  input hospital $ treatment $ events trials;
  cards;
  low treat 5 104
  low cont 10 105
  med treat 35 110
  med cont 102 197
  high treat 95 105
  high cont 99 104
  ;
run;
  
proc logistic data = hosp; 
  class hospital (ref = 'low')
        treatment (ref = 'cont') / param = ref; 
  model events/trials = treatment hospital / cl; 
run; 
```

### Individual level data

```{r, engine = "sas", engine.path = "C:/Program Files/SASHome/SASFoundation/9.4/sas.exe"}
title 'Example Low Birth-weight and Maternal Age'; 
proc import datafile = 'C:\Users\niwi8\OneDrive\Documents\p8120_ta\data\lowbwt.csv'
            out = lowbwt
            dbms = csv; 
run; 
            
proc logistic data = lowbwt descending; 
  model low = age / waldrl; 
  units age = 5; 
run; 
```

### Likelihood ratio test

```{r, engine = "sas", engine.path = "C:/Program Files/SASHome/SASFoundation/9.4/sas.exe"}
proc import datafile = 'C:\Users\niwi8\OneDrive\Documents\p8120_ta\data\lowbwt.csv'
            out = lowbwt
            dbms = csv; 
run; 

title1 'Likeliohood ratio test';
title2 'LNull/Reduced Model'; 
proc logistic data = lowbwt descending; 
  class smoke (ref = '0') / param = ref; 
  model low = smoke age / cl; 
run; 

title2 'Full Model'; 
proc logistic data = lowbwt descending; 
  class smoke (ref = '0')
        visits (ref = '0') / param = ref; 
  model low = smoke age visits / cl; 
run; 
```








