---
title: "Likelihood ratio test and 2-sample test of proportions"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

The notes for the topics on this page can be found in the lectures 9 folders on Canvas. 

# {.tabset .tabset-fade}

## R

To conduct a likelihood ratio test on a 2 x 2 table in R you can use `GTest()` from the *DescTools* package. To conduct a 2-sample test of proportions you can use `prop_test()` from the *catfun* package. For both functions you can provide a frequency table as the main argument as seen in the examples below. 

```{r}
vietnam <- tibble(
  service = c(rep("yes", 2), rep("no", 2)), 
  sleep = c(rep(c("yes", "no"), 2)), 
  count = c(173, 160, 599, 851)
)

vietnam <- vietnam %>% 
  mutate(service = as_factor(service), 
         sleep = as_factor(sleep))
```

### Likelihood ratio test

```{r}
library(DescTools)

xtabs(count ~ service + sleep, data = vietnam) %>% 
  GTest()
```

### 2-sample test of proportions

```{r}
library(catfun)

xtabs(count ~ service + sleep, data = vietnam) %>% 
  prop_test()
```

## SAS

To conduct a likelihood ratio test for a frequency table and a two-sample test of proportions in SAS you use the frequency procedure. To indicate that you want SAS to return a likelihood ratio test you specify the `chisq` argument in the options of the `table` statement. To indicate you want SAS to return a two-sample test of proportions you can specify `riskdiff(equal var = null)` in the options of the `table` statement.  

### Likelihood ratio test

```{r, engine = "sas", engine.path = "C:/Program Files/SASHome/SASFoundation/9.4/sas.exe"}
data vietnam; 
  input service $ sleep $ count; 
  cards; 
  yes yes 173
  yes no 160
  no yes 599
  no no 851
  ; 
run; 

proc freq data = vietnam order = data; 
  table service*sleep / relrisk chisq expected; 
  weight count; 
run; 
```

### 2-sample test of proportions

```{r, engine = "sas", engine.path = "C:/Program Files/SASHome/SASFoundation/9.4/sas.exe"}
data vietnam; 
  input service $ sleep $ count; 
  cards; 
  yes yes 173
  yes no 160
  no yes 599
  no no 851
  ; 
run; 
  
proc freq data = vietnam order = data; 
  table service*sleep / expected riskdiff(equal var = null); 
  weight count; 
run; 
```





