---
title: "Measures of Association"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

The notes for the topics on this page can be found in the lectures 5 and 7 folders on Canvas. 

# {.tabset .tabset-fade}

## R

Odds ratios and risk ratios can be calculated from contingency tables using the functions `oddsratio()` and `riskratio()` from the *epitools* package. Risk difference can be calculated from contingency tables using the `riskdiff()` function from the *catfun* package. For further help, please watch the video [example](#r-demonstration).

```{r warning=FALSE}
library(epitools)
library(catfun)

trial <- tibble(
  disease = c(rep("yes", 2), rep("no", 2)), 
  treatment = c(rep(c("estrogen", "placebo"), 2)), 
  count = c(751, 623, 7755, 7479)
)

trial <- trial %>% 
  mutate(disease = as.factor(disease), 
         treatment = as.factor(treatment),
         disease = fct_relevel(disease, "yes", "no"), 
         treatment = fct_relevel(treatment, "estrogen", "placebo"))
```

For these examples, the `oddsratio()` and `riskratio()` are used in combination with `xtabs()` while `riskdiff()` only calls upon the dataset. 

### Odds ratio

```{r}
xtabs(count ~ treatment + disease, data = trial) %>% 
  oddsratio(method = "wald")
```

### Risk ratio

```{r}
xtabs(count ~ treatment + disease, data = trial) %>% 
  riskratio()
```

### Risk difference

```{r}
trial %>% 
  riskdiff("treatment", "disease", weight = "count")
```

### R demonstration

<div align="center">
<iframe width="854" height="480" src="https://www.youtube.com/embed/LWvuOJ986xI" frameborder="0" allowfullscreen></iframe>
</div>

## SAS

To caculate measures of association from a contingency table in SAS you use the frequency procedure with the `relrisk` argument in table statement to generate the relative risk/odds ratio and the `riskdiff` argument to generate the risk difference. This will always produce confidence intervals for generated measures. The syntax is: 

```{}
proc freq data = <dataset>; 
  table <row variable>*<column variable> / <relrisk> <riskdiff>; 
  weight <weighting variable if summary data>; 
run; 
```

```{r, engine = "sas", engine.path = "C:/Program Files/SASHome/SASFoundation/9.4/sas.exe"}
data trial; 
  input disease $ treatment $ count; 
  cards; 
  yes estrogen 751
  yes placebo 623
  no estrogen 7755
  no placebo 7479
  ; 
run; 
  
proc freq data = trial order = data; 
  table treatment*disease / riskdiff relrisk; 
  weight count; 
run; 
```


