---
title: "Measures of Association"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

The notes for the topics on this page can be found in the lecture 5 folder on Canvas. 

# {.tabset .tabset-fade}

## R

Odds ratios and risk ratios can be calculated from contigency tables using the functions `oddsratio()` and `riskratio()` from the *epitools* package. Risk difference can be calculated from contingency tables using the `riskdiff()` function from the *catfun* package.

```{r warning=FALSE}
library(epitools)
library(catfun)

trial <- tibble(
  disease = c(rep("yes", 2), rep("no", 2)), 
  treatment = c(rep(c("estrogen", "placebo"), 2)), 
  count = c(751, 623, 7755, 7479)
)

trial <- trial %>% 
  mutate(disease = as.factor(disease), 
         treatment = as.factor(treatment),
         disease = fct_relevel(disease, "yes", "no"), 
         treatment = fct_relevel(treatment, "estrogen", "placebo"))
```

### Odds ratio

```{r}
xtabs(count ~ treatment + disease, data = trial) %>% 
  oddsratio(method = "wald")
```

### Risk ratio

```{r}
xtabs(count ~ treatment + disease, data = trial) %>% 
  riskratio()
```

### Risk difference

```{r}
trial %>% 
  riskdiff("treatment", "disease", weight = "count")
```

## SAS

```{r, engine = "sas", engine.path = "C:/Program Files/SASHome/SASFoundation/9.4/sas.exe"}
data trial; 
  input disease $ treatment $ count; 
  cards; 
  yes estrogen 751
  yes placebo 623
  no estrogen 7755
  no placebo 7479
  ; 
run; 
  
proc freq data = trial order = data; 
  table treatment*disease / riskdiff relrisk; 
  weight count; 
run; 
```


