---
title: "Power"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(SASmarkdown)
```

# {.tabset .tabset-fade}

## R

The *pwr* package has a variety of functions for conducting power analyses. It is important to note, however, that these functions to not return the exact same results as SAS. 

### Computing power for fixed sample size

The `pwr.2p.test()` function can be used to calculate power for equal group sample sizes. The first arument `h` takes Cohen's H as the measure of effect; to calculate Cohen's H use the `ES.h()` function as demonstrated. Unlike SAS, the `n` argument is the number of observations in each group and not the entire sample. 

```{r warning = FALSE}
library(pwr)

pwr.2p.test(h = ES.h(0.35, 0.2), 
            n = 110, 
            sig.level = 0.05, 
            alternative = "two.sided")
```

### Computing sample size given power (balanced)

The `pwr.2p.test()` function can also be used to calculate sample size given power. Instead of specifying the number of observations per group, we specify the desired power. 

```{r}
pwr.2p.test(h = ES.h(0.35, 0.2), 
            power = 0.80, 
            sig.level = 0.05, 
            alternative = "two.sided")
```

### Computing sample size given power (unbalanced)

TBD

## SAS

Power analyses in SAS are conducting using the power procedure. 

### Computing power for fixed sample size

```{sas, engine.path = "C:/Program Files/SASHome/SASFoundation/9.4/sas.exe", error = FALSE}
title1 'Computing Power Given Sample Size';
proc power; 
  twosamplefreq test = pchi
  alpha = 0.05
  groupproportions = (0.35 0.20)
  nullproportiondiff = 0
  groupweights = (1 1)
  ntotal = 220
  power = .;
run;
```

### Computing sample size given power (balanced)

```{sas, engine.path = "C:/Program Files/SASHome/SASFoundation/9.4/sas.exe", error = FALSE}
title1 'Computing Sample Size Given Power';
proc power; 
  twosamplefreq test = pchi
  alpha = 0.05
  groupproportions = (0.35 0.20)
  nullproportiondiff = 0
  npergroup = .
  power = 0.80;
run;
```

### Computing sample size given power (unbalanced)

```{sas, engine.path = "C:/Program Files/SASHome/SASFoundation/9.4/sas.exe", error = FALSE}
title1 'Computing Sample Size Given Power (Unbalanced Groups)';
proc power; 
  twosamplefreq test = pchi
  alpha = 0.05
  groupproportions = (0.35 0.20)
  nullproportiondiff = 0
  groupweights = (2 1)
  ntotal = .
  power = 0.85;
run;
```


