---
title: "R"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(epitools)
library(DescTools)
library(descr)
library(survival)
library(catfun)

data(esoph)

azt <- read_csv("../data/AZT.csv")

azt <- azt %>% 
  mutate(aids = as.factor(aids), 
         aids = fct_relevel(aids, "1", "0"), 
         azt = as.factor(azt), 
         azt = fct_relevel(azt, "1", "0"))
```

```{r, echo = FALSE, out.width = "10%"}
htmltools::img(src = knitr::image_uri("../logos/p8120_logo.png"), 
               alt = 'logo', 
               style = 'position:absolute; top:0; right:300px; padding:10px; width:125px; height:125px;')
```

<br>

The R functions within this document come from a variety of libraries. To use these functions please intsall the following packages by running: 

`install.packages(c("tidyverse", "descr", "descTools", "surival", "epitools"))` 

In addition, you will need to run, `devtools::install_github("nt-williams/catfun")`

### Cross-tabulation

To create sparse cross-tabulations use the function `xtabs()`

```{r xtabs example}
# data(esoph)
# 
# xtabs(ncontrols ~ alcgp + agegp, data = esoph)
# 
# xtabs(~ aids + azt, data = azt) %>% 
#   oddsratio(method = "wald") 
# 
# xtabs(~ aids + azt, data = azt) %>% 
#   riskratio(method = "wald")
```


### Hypothesis tests

Use the following table to locate functions for hypothesis tests. If you are unsure of the usage of a function type `?<insert function>`

|Hypothesis test                     | R function                     | Package          |
|------------------------------------|--------------------------------|------------------|
|One/two sample test proportions, normal approximation  |`prop.test()`| stats            |
|One/two sample test proportions, exact   | `binom.test()`            | stats            |
|Confidence intervals, proportions   | `BinomCI()`                    | DescTools        |
|Chi-squared                         | `crosstab(..., chisq = TRUE)` | descr             |
|Fisher's exact                      | `crosstab(..., fisher = TRUE)`| descr             |
|McNemar's test                      | `crosstab(..., mcnemar = TRUE)`   | descr         |
|Logistic regression                 | `glm(..., family = binomial)`  | stats            |
|GOF & Hosmer Lemeshow               | TBD                            | TBD              |
|Conditional logistic regression     | `clogit()`                     | survival         |
|Cochran Armitage Trend test         | `CocrhanArmitageTest()`        | DescTools        |
|Brewlow Day test                    | `BreslowDayTest`               | DescTools        |

### Measures of association

| Measure               | R function         | Package
|-----------------------|--------------------|----------
|Odds ratio (2xk table) | `oddsratio()`      | epitools
|Risk ratio (2xk table) | `riskratio()`      | epitools
|Risk difference        | `riskdiff()`       | catfun

### Power

Fun

